\setcounter{secnumdepth}{0}
\section{Introduction}
This document analyses and simulates the behavior of a digital transmission system to have a better understanding of the concept behind these types of telecommunication. The document is split into two main parts:

\begin{enumerate}
    \item[\texttt{I}] \hspace{10px} \textbf{Analysis.} The first part of the document analyzes the transmission system properties following the tasks of the course project guidelines. The numerical outcome of these tasks is summarized in the \hyperref[raw-results]{raw results} at the end of the analysis part of the document.
    % 
    \item[\texttt{II}] \hspace{10px} \textbf{Simulation.} The second part of the document focuses on the functions used to properly simulate a digital transmission system. In the simulation part, there will be the explained MATLAB code produced to perform all the processes described in the \hyperref[general-schematic]{general schematic}.
\end{enumerate}



\noindent The full project can be found and downloaded on the public GitHub repository \texttt{\href{https://github.com/imAlessas/transmission-simulation.git}{imAlessas/transmission-simulation.git}} where it is possible to find the full \href{https://github.com/imAlessas/transmission-simulation/tree/main/src}{MATLAB code} of the project and the \href{https://github.com/imAlessas/transmission-simulation/tree/main/doc}{\LaTeX\, code} of the documentation.


\subsection{General schematic}\label{general-schematic}
The full schematic - containing every step - of a transmission system is presented in figure\,\ref{fig:transmission_diagram}. Before exploring the mathematical background hidden between the steps, it is crucial to understand what every phase of the system means.

\begin{itemize}
    \renewcommand{\labelitemi}{$\diamond$}
    \item \textsl{Source.} The source device is whichever device is sending a signal; it could be a television, a computer, a smartphone, or anything else.
    % 
    \item \textsl{Formatting Device.} The formatting device's task is to translate the information from analogic to digital which translates into sampling the continuous analogic signal and creating a discrete digital signal that can be transmitted through digital devices.
    % 
    \item \textsl{Source Coding.} The source coding goal is lossless data compression. Sure enough, through the Shannon-Fano source coding, the symbols transmitted are encoded to reduce the average codeword length.
    % 
    \item \textsl{Channel Coding.} The channel coding goal is to add some control bits that will help detect and eventually correct the errors that occurred during the transmission.
    % 
    \item \textsl{Interleaving.} The interleaver is needed to transform package errors into independent errors. This is achieved by changing the ordering of the symbols that will be transmitted.
    % 
    \item \textsl{Scrambling.} The scrambling procedure helps with the synchronization between the two devices and improves the security of the transmission. This is achieved by adding a \textit{pseudo-random sequence} to the symbols before the transmission.
    % 
    \item \textsl{Modulation.} The modulation process' goal is to match the spectrum of the transmitted signal with the transmission channel bandwidth making the signal more noise-immune and increasing the data-transfer rate; these operations are performed by the modulator. There are different types of modulation, the one utilized in this project is the \textsl{Binary Phase Shift Keying}, which is one of the most effective modulations against noise. 
    % 
    % 
    \input{../res/TikZ/diagram.tex}
    % 
    % 
    \item \textsl{Noise.} The noise is a crucial obstacle to overcome to have a successful transmission; the noise is the main reason for a wrongly transmitted symbol. There are different types of noise, some of them are generated by other transmissions, others are due to the physical medium and others are caused by the intermediate devices between the transmission. Nevertheless, in every transmission, there will be the \textsl{Gaussiam White Noise} which is a thermal noise caused by the Big Bang.
    % 
    \item \textsl{Demodulation.} In this phase the demodulator device, after receiving the disturbed signal, will try to detect the signal to regenerate the original one. Sometimes the noise energy will be stronger than the signal energy generating errors that will be corrected in the next steps.
    % 
    \item \textsl{Descrambling.} The descrambling procedure is the opposite of the scrambling. The added \textit{pseudo-random sequence}, after the reception is subtracted by the descrambler.
    % 
    \item \textsl{Deinterleaving.} The deinterleaver reorders the transmitted symbols in the opposite way that the interleaver did. In such a way the \textit{burst} errors that occurred during the transmission will become single errors that can be easily recovered.
    % 
    \item \textsl{Channel Decoding.} The channel decoding process uses the added bits during the channel encoding to perform an error correction algorithm that will drastically decrease the error rate of the transmission.
    % 
    \item \textsl{Source Decoding.} The source decoding procedure decompresses the received data into the original symbols. This is achieved by one of the source coding properties: symbols are easily detected because there are no shorter codes at the beginning of longer codes.
    % 
    \item \textsl{Formatting Device.} During the transmission this device converts the signal from analogic to digital, during the reception of the signal the formatting device translates the discrete digital signal into a continuous analogic signal. 
    % 
    \item \textsl{Destination.} The destination device is whichever device will receive the signal. Likewise the source one, the destination device could be a satellite, a smartphone, a server, or anything else.
\end{itemize}

\subsection{Initial parameters}\label{initial-parameters}
The parameters used in this project have been assigned in a datasheet and are reported in the following list:
\vspace{-5px}
\begin{itemize}
    \renewcommand{\labelitemi}{$\cdot$}
    \setlength{\itemsep}{-2px}
    \item \textsl{Symbol duration}: 60 ns, also called $\tau$;
    \item \textsl{SNR}: 8.1 dB;
    \item \textsl{Source code}: Shannon-Fano coding;
    \item \textsl{Error correction code}: cyclic coding with codeword length \textit{m = 31} and generator polynomial $ z^5 \oplus z^2 \oplus 1$;
    \item \textsl{Carrier frequency}: 2.5 GHz;
    \item \textsl{Modulation}: Binary Amplitude Shift Keying (BPSK) with the phase shift of $\pi$.
\end{itemize}

\noindent In addition, the source data (alphabet) and the symbols' respective probabilities are summarized in the following table.

\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|}
        \hline
        \multicolumn{2}{|c|}{\texttt{Source 7}} \\\hline\hline
        $ a_1 $ & 0.11 \\
        $ a_2 $ & 0.07 \\
        $ a_3 $ & 0.09 \\
        $ a_4 $ & 0.01 \\
        $ a_5 $ & 0.06 \\
        $ a_6 $ & 0.06 \\
        $ a_7 $ & 0.13 \\
        $ a_8 $ & 0.14 \\
        $ a_9 $ & 0.13 \\
        $ a_{10} $ & 0.05 \\
        $ a_{11} $ & 0.11 \\
        $ a_{12} $ & 0.04 \\\hline
    \end{tabular}
    \label{tab:source7}
\end{table}

\FloatBarrier\noindent Afterwards the parameters have been transcripted in the MATLAB program. In the following code snippet, other than the initial parameters, some other useful values have been calculated such as the number of symbols to be transmitted, the number of samples per symbol, the $r$ and $k$ values which play a crucial role in the code correction algorithms (see\,\ref{ciclic-coding},\ref{hamming-encoding}) and finally the scrambler key (see\,\ref{scrambling}).

\begin{lstlisting}
    % Initial constant parameters

    % source number 7
    alphabet = [1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12];
    probability_vector = [11, 7, 9, 1, 6, 6, 13, 14, 13, 5, 11, 4]/100;
    
    tau = 60e-9;                        % symbol duration time, [s]
    SNR = 8.1;                          % Signal-to-Noise-Ration, [dB]
                                        % Source Code: Shannon-Fano
                                        % Error correction code: Cyclic
    generation_polynomial = [...        % z^5 + z^2 + z^0
        1 0 0 1 0 1]; 
    codeword_length = 31;               % m
    
    f0 = 2.5e+9;                        % carrier frequency [Hz]
                                        % Modulation: BPSK
    phase_shift = pi;                   % Phase shift [rad]
    U = 1;                              % amplitude BPSK signal [V]
    
    % Additional data
    transmitted_symbol_number = randi(1e5,1);     % number of symbols   
    samples_per_symbol = 500;           % samples per symbol
    r = ceil(log2(codeword_length + 1));% r
    k = codeword_length - r;            % k
    
    % Used for the scrambling/descrmabling algorithm
    scrambler_key = randi(2, 1, codeword_length) - 1; 
\end{lstlisting}





\setcounter{secnumdepth}{1}

